An Efficient Calculus for Generalised Parikh Images of Regular Languages

### 2022-06-06 main.tex
Veckans text tm

\usepackage{amsthm}
\usepackage{amsmath,mathtools,empheq,fancybox}
\usepackage{paralist}
\usepackage{url}
\usepackage{color}
\usepackage{textcomp,listings}
\usepackage{array}
\usepackage{mymacros}
\usepackage{microtype}
\usepackage{listings}
\usepackage{csquotes}
\usepackage{proof}

@@

Addressing these concerns, our specialised calculus for Parikh images of
products of regular languages allows us to interleave the computation of
arbitrarily deep products of automata with the product's Parikh image. This
allows us to let both calculations inform each other, eliminating unnecessary
work, and pruning the size of the partial products considered in the
computation. Moreover, the method can be used iteratively to tackle smaller
chunks of the product incrementally, thereby avoiding running out of memory.

\subsection{Motivating Example}

@@

\begin{equation}
\begin{aligned}
\psi(\mathcal{A}) := 
& \bigwedge_{\alpha \in \Sigma}
c_\alpha = \sum_{\delta \in \delta} t_\delta  
\text{ where $\alpha \in \delta$}\\
&\bigwedge_{q \in Q} \left (\text{$1$ if $q \in I$} +
\sum_{\delta = q' \xrightarrow{} q} t_\delta 
- \sum_{\delta = q\xrightarrow{}q'} t_\delta \right)
\begin{cases}
\geq 0 & \text{if $q \in I$} \\
= 0 & \text{ otherwise}
\end{cases}\\
& \bigwedge_{\delta = q \xrightarrow{} q'} t_\delta > 0 
\implies z_{'q} > 0 \\
& \bigwedge_{q, q' \in Q} z_{q'} > 0 
\implies 
\begin{cases}
\bigvee\limits_{\delta = q \xrightarrow{} q'} z_{q'} = z_{q} + 1 \land t_\delta > 0 \land z_{q} > 0 & \text{if $q \not\in  I$} \\
z_{q'} = 1 \land t_\delta > 0& \text{if $q \in I$}
\end{cases}
\end{aligned}
\label{eq:generate-parikh}
\end{equation}

where all variables $z_i, t_i$ are existentially quantified and the variables
$c_\alpha$ make up the actual image. Feeding this definition into a Presburger
solver and performing quantifier elimination on $z_i, t_i$ would produce the
Parikh image in its Presburger form.

@@

\subsubsection{An Example}

\Fudge{AN EXAMPLE!}

\section{A Calculus for Generalised Parikh Images}

@@

@@

\subsection{Lazy Computation of Parikh Images for Regular Languages}

@@

\section{Backjumping and Learning No-Goods}

@@

We implement the baseline approach using the same Presburger solver
(\Princess{}), input file parser, and automaton implementation as \Catra{}. We
do this in order to make the comparison between methods as fair as possible, and
give us the ability to compare the efficiency of the calculus rules themselves.
Using the formula of Equation~\ref{eq:generate-parikh}, we produce quantified
Presburger formulae for each successive term and add them to Princess. In this
fashion we compute the product incrementally term by term, checking
satisfiability in each step. We use \Fudge{a priority queue} to select which
term to use for the next step of the product, and order it by \Fudge{the number
of transitions} as a heuristic for the size of the term. We use this heuristic
to avoid computing large (and therefore slow) products until we have to, banking
our hopes on finding a source of unsatisfiability earlier.

@@

\subsection{Tuning and optimising our Implementation}
\contents{
  \item Optimisations and tricks
}

@@

@@

\section{Threats to Validity}
\contents{
  \item There might be better implementations
  \item We might have over-biased
  \item Our benchmarks might be biased
  \item The underlying automata implementation might be bad and therefore skew
  the results.

@@

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

>>>>

\usepackage{amsthm}
\usepackage{amsmath,mathtools,empheq,fancybox}
\usepackage{paralist}
\usepackage{url}
\usepackage{color}
\usepackage{textcomp,listings}
\usepackage{array}
\usepackage{mymacros}
\usepackage{microtype}
\usepackage{listings}
\usepackage{csquotes}
\usepackage{proof}
\usepackage{algorithm2e}

@@

Addressing these concerns, our calculus for Parikh images of products of regular
languages allows us to interleave the computation of arbitrarily deep products
of automata with the product's Parikh image. This allows us to let both
calculations inform each other, eliminating unnecessary work, and pruning the
size of the partial products considered in the computation. Moreover, the method
can be used iteratively to tackle smaller chunks of the product incrementally,
thereby avoiding running out of memory.

\subsection{A demonstration of our approach}

@@

\begin{equation}
\begin{aligned}
\ParikhMap(\Automaton) := 
& \AndComp{\Letter \in \Alphabet}{\LetterVar_{\Letter} = \sum_{\Transition = \FromLabelTo{}{\Letter}{} \in \Transitions} \TransitionVar_{\Transition}
}\\
& \AndComp{\State \in \AcceptingStates}{\FinalStateVar_{\State} > 0 \implies \StateVar_{\State} > 0}\\
& \AndComp{\State \in \States}{
  \left(\FinalStateVar_{\State} \text{ if } \State \in \AcceptingStates \text{ otherwise } 0 \right) +
  \sum_{\Transition = \FromLabelTo{\State}{}{}} - \sum_{\Transition = \FromLabelTo{}{}{\State}}
= \begin{cases}
    1 \text{  if $\State = \InitialState$} \\
    0 \text{ otherwise}
  \end{cases}
}\\
& \AndComp{\Transition = \FromLabelTo{}{}{\State} \in \Transitions}{
  \TransitionVar_{\Transition} > 0 \implies \StateVar_{\State}
} \\
& \AndComp{\State \in \States}{
  \StateVar_{\State} > 0 \implies
  \left(\FinalStateVar_{\State} = 1 \text{ if } \State \in \AcceptingStates\right) \OrComp{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions}{
    \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
    \TransitionVar_{\Transition} \geq 1 \land
  \StateVar_{\State'} \geq 1
    }
}
\end{aligned}
\label{eq:generate-parikh}
\end{equation}

where all variables $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ are existentially quantified and the variables
$\LetterVar_\Letter$ make up the actual image. Feeding this definition into a Presburger
solver and performing quantifier elimination on $\TransitionVar_i, \StateVar_i, \FinalStateVar_i$ with $\LetterVar_\Letter$ as free variables would produce the
Parikh image in its Presburger form.

@@

\subsubsection{An Example}
We apply Equation~\ref{eq:generate-parikh} to the automaton of Figure~\ref{fig:len-branch}. We then get the following quantified Presburger formula. As before, we omit the implicit existential quantifiers for brevity, along with the constranints requiring all variables to be nonnegative. We also omit the $\land$-clauses between each line. In other words, in this example, only the variables $\LetterVar_a, \LetterVar_b, \LetterVar_c, \LetterVar_d, \LetterVar_e$, and $\LetterVar_f$ are free, and corresponds to the counts of the respective subscript letter. Note that since they here uniquely identify a transition due to the one-to-one mapping between character labels and transitions in the example, the first clause can be trivially eliminated by substituting $\LetterVar_a$ for $\TransitionVar_{\FromLabelTo{0}{a}{1}}$, etc. This is not generally the case.

\begin{equation}
  \begin{aligned}
  \ParikhMap(\Automaton):=\:
  &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} \land \LetterVar_{d} = \TransitionVar_{\FromLabelTo{2}{d}{3}} \land \LetterVar_{e} = \TransitionVar_{\FromLabelTo{2}{e}{2}} \land \LetterVar_{f} = \TransitionVar_{\FromLabelTo{1}{f}{3}}\\
  &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
  & \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{0}{a}{1}} - 0 = 1 \land  \TransitionVar_{\FromLabelTo{1}{f}{3}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
  & \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3} + 0 - \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{1}{f}{3}} = 0\\
  &\TransitionVar_{\FromLabelTo{0}{a}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} > 0 \implies \StateVar_{2} > 0\\
  &\TransitionVar_{\FromLabelTo{1}{c}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{2}{e}{2}} > 0 \implies \StateVar_{2} > 0\\
  &\TransitionVar_{\FromLabelTo{1}{f}{3}} > 0 \implies \StateVar_{3} > 0 \land \TransitionVar_{\FromLabelTo{2}{d}{3}} > 0 \implies \StateVar_{3} > 0\\
  &\StateVar_{0} > 0 \implies \left(\StateVar_{0} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{0}{a}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{0} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} \geq 1 \land \StateVar_{2} \geq 1\right)\\
  &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{1}{c}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{1} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{1}{f}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
  &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{2}{e}{2}} \geq 1 \land \StateVar_{2} \geq 1\right) \lor \left(\StateVar_{2} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{2}{d}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
  &\StateVar_{3} > 0 \implies \FinalStateVar_{3} = 1
  \end{aligned}
  \end{equation}

\section{Overview of our generalised Parikh images}

@@

\subsection{Applications}

@@

\section{A Calculus for Generalised Parikh Images}\label{sec:calculus}

@@

\section{Extensions}
\subsection{Backjumping and Learning No-Goods}
\subsubsection{An Example}
\subsection{Symbolic Automata}\label{sec:ext:symbolic}

@@

We implement the baseline approach using the same Presburger solver
(\Princess{}), input file parser, and automaton implementation as \Catra{}. We
do this in order to make the comparison between methods as fair as possible, and
give us the ability to compare the efficiency of the calculus rules themselves.
Using the formula of Equation~\ref{eq:generate-parikh}, we produce quantified
Presburger formulae for each successive term and add them to Princess. In this
fashion we compute the product incrementally term by term, checking
satisfiability in each step. We use \Fudge{a priority queue} to select which
term to use for the next step of the product, and order it by \Fudge{the number
of transitions} as a heuristic for the size of the term. We use this heuristic
to avoid computing large (and therefore slow) products until we have to, banking
our hopes on finding a source of unsatisfiability earlier. The pseudocode for
our implementation can be seen in Algorithm~\ref{alg:baseline}.

\begin{algorithm}
  \caption{An algorithm with caption}\label{alg:baseline}
  \KwData{$\Automaton_1, \ldots, \Automaton_n$ automata, other constraints $\SomeClause$}
  \KwResult{\textsc{Sat} or \textsc{Unsat}}
  \SetKwFunction{NewTheoremProver}{newTheoremProver}
  \SetKwFunction{NewPriorityQueue}{newPriorityQueue}
  \SetKwFunction{Dequeue}{dequeue}
  \SetKwFunction{Enqueue}{enqueue}
  \SetKwFunction{Assert}{assert}

  $p \gets \NewTheoremProver{}$

  \Assert{$p$, $\SomeClause$}

  \ForEach{$\Automaton_i$}{
    \Assert{$p, \ParikhMap(\Automaton_i)$}

    \If{$p$ is \textsc{Unsat}}{break}

  $q \gets \NewPriorityQueue{}$

  \While{$p$ not \textsc{Unsat} and $|q| > 1$}{
    $\Automaton, \Automaton' \gets \Dequeue{q}$

    \Assert{$p, \ParikhMap(\Automaton \times \Automaton')$}

    \Enqueue{$q, \Automaton \times \Automaton'$}
  }

  \KwRet{$p$'s SAT status}

  \end{algorithm}

@@

\subsection{Heuristics and search strategies}
\contents{
  \item Optimisations and tricks
}

@@

\subsection{Threats to Validity}
\contents{
  \item There might be better implementations
  \item We might have over-biased
  \item Our benchmarks might be biased
  \item The underlying automata implementation might be bad and therefore skew
  the results.

@@

@@

%% Appendix
\appendix
% \section{Appendix}

% Text of appendix \ldots
